---
title: "Peer Assessment 1"
author: "Evandro Armelin"
date: "8 de novembro de 2015"
output: html_document
---

This is a R Markdown document that describes the resolution and results of the Peer Assessment 1 of Reproducible Research course from Coursera

The next sections of the documents will show the code chuncks and results of each analysis required by the Assessment

##1. Loading and Processing the Data

```{r, message=F, warning=F}
#loading libraries
library(Hmisc)
library(ggplot2)
library(timeDate)

#define the working directory
setwd("C:/Users/Evandro/Note_pessoal/Documentos - note pessoal/Evandro/Didáticos/Data Scientist/05. Reproducible Research/Week 2/Course Project 1")

#read the original data
data <- read.csv("repdata_data_activity/activity.csv")
```


##2. What is mean total number of steps taken per day?

###Histogram of Steps per Day
```{r}
#aggregate steps per day
daysteps <- aggregate(data$steps, by = list(date=data$date), FUN = sum, na.rm = TRUE)

#adjust variables names
names(daysteps) <- c("Date", "Steps")

hist(daysteps$Steps,
     main = "Histogram of Steps per Day",
     xlab = "Steps per Day",
     col = "lightblue",
     breaks = 10)
```

###Mean and median calculation
```{r}
mean(daysteps$Steps)
median(daysteps$Steps)
```


##3. What is the average daily activity pattern?

###Calculating interval means through all the day
```{r}
intsteps <- aggregate(data$steps, by = list(data$interval), FUN = mean, na.rm = TRUE)
names(intsteps) <- c("Interval", "Average_Steps")
```

###Plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days
```{r}
plot(intsteps$Interval,
     intsteps$Average_Steps, type = "l",
     main = "Average Steps for each Interval (across all the days)",
     xlab = "Interval",
     ylab = "Average Stepes",
     col = "Red")
```

###Calculate the Interval with the maximun average of steps
```{r}
intsteps[which.max(intsteps$Average_Steps),]
```

##4. Imputing missing values

###Calculate the total number of missing values in the dataset
```{r}
sum(!complete.cases(data))
```

###Imputing data for the NAs data in a new dataset equal to the original one
```{r}
completedata <- data
completedata$steps <- impute(data$steps, fun = mean)
```

###Histogram of Steps per Day for the Completed Data
```{r}
#aggregate steps per day
completedatadaysteps <- aggregate(completedata$steps, by = list(date=completedata$date), FUN = sum, na.rm = TRUE)

#adjust variables names
names(completedatadaysteps) <- c("Date", "Steps")

hist(completedatadaysteps$Steps,
     main = "Histogram of Steps per Day (completed data)",
     xlab = "Steps per Day",
     col = "lightgreen",
     breaks = 10)
```

###Mean and median calculation for the Completed Data
```{r}
mean <- c(mean(completedatadaysteps$Steps), mean(daysteps$Steps))
median <- c(median(completedatadaysteps$Steps), median(daysteps$Steps))
comp <- data.frame(rbind(mean, median))
names(comp) <- c("Completed Data", "Original Data")
comp

```

##5. Differences in activity patterns between weekdays and weekends

###Calculating interval means through all the days, spliting in Weekday and Weekend
```{r}
#Creating factor variable to set Weekend and Weekday
data$Weekday <- factor(c("Weekday", "Weekend"))
data[isWeekend(data$date) == TRUE,4] <- "Weekend"
data[isWeekend(data$date) != TRUE,4] <- "Weekday"

intstepsweekday <- aggregate(data$steps, by = list(data$interval, data$Weekday), FUN = mean, na.rm = TRUE)
names(intstepsweekday) <- c("Interval", "Weekday", "Average_Steps")
```

###Plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days, splited by Weekend and Weekday
```{r}
qplot(data = intstepsweekday, Interval, Average_Steps, facets = Weekday~., geom = "line", color = Weekday, ylab = "Average Steps")